A Novel Model-Free Learning Approach for Dynamic Target Tracking Problem and its Validation using Virtual Robot Experimentation Platform

### 2020-06-28 paperMIC2020/paperIEEE-MIC2020-V1.tex

 \section{Computer Experiments and Results}
 \label{sec:resultsExperiments}
 In the sequel, computer experiments  are conducted to validate the performance of the proposed model-free adaptive learning algorithm in real-time. The results of computer experiments highlight the dynamics of the tracking errors and the convergence characteristics of the proposed algorithm (i.e., updating the critic weights). This will judge the ability of the follower robot to track independent motion trajectory of the leader robot under two different independent leader-motion trajectories. The computer experiments are realized using MATLAB simulation environment. The weighting matrices are set to $\mathbf{Q} = \mathrm{diag}[0.01,0.01,0.005]$ and $\mathbf{R} = \mathrm{diag}[10^{-6}, 10^{-6}].$ % 
  \[Q=0.001 
  \begin{bmatrix}
  10       & 0   & 0\\
  0       & 10   & 0\\
  0       & 0   &5 \\
  \end{bmatrix},
  R=0.000001
  \begin{bmatrix}
  1       & 0 \\
  0       & 1 
  \end{bmatrix}.\]
 The learning rate $\ell_c$ is set to  $0.0001$. The sampling time $T_s$ is set to $0.08 \sec.$ The desired distance offset between the leader and the follower is set to $ d = 0.5$ [m] for all scenarios.

% In the first scenario, the leader trajectory motion is described by a unicycle model so that %

%\begin{align*}
% $\dot x^{[\ell]}(t) = \nu^{[\ell]}(t)\cos\theta^{[\ell]}(t),$ %
% $\dot y^{[\ell]}(t) = \nu^{[\ell]}(t)\sin\theta^{[\ell]}(t).$
% \dot\theta^\ell(t) &= \omega(t),
%\end{align*}

 Initially, the leader's position is set to be the same for a short period of time with $\nu^{\ell} = 0 , \gamma^{\ell} = 0$. Then it starts to move on a horizontal line with $\nu^{\ell} = 0.1 $ [m/s], eventually the leader starts to move on an inclined path with $\nu^{\ell} = 0.2$ [m/s], $\gamma^{\ell} = 45\degree$. The leader is initially placed at $(x,y) = (5, 2)~\si{[\meter]}$ with an orientation of $\theta = 0 \degree $, while the follower  is initially placed at $(x,y) = (0, 0)~\si{[\meter]}$ with an orientation of $\theta = 0\degree$. Note that during this scenario, the critic weights converge rapidly and the tracking errors decay till the follower tracks the leader. The trajectory phase plan plot, tracking error states, control signals, and tuning of critic weights are shown in Fig.~\ref{fig:performanceRectilinearTrajectory}. %

% \begin{figure*}[htbp]%
%     \centering
%     \parbox[c]{0.48\textwidth}{\centering Parallel parking}
%     \parbox[c]{0.48\textwidth}{\centering Trajectory tracking}
%     \\
%     \subfigure[][]{%
%         \label{fig:trajectoryRectilinearDistance}%
%         \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/matlab/rectilinear/trajectoryRectlinearDistance} }
%     \subfigure[][]{%
%         \label{fig:stateErrorRectilinearDistance}%
%         \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/matlab/rectilinear/stateErrorRectlinearDistance} }
%     \\
%     \subfigure[][]{%
%         \label{fig:controlInputsRectilinearDistance}%
%         \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/matlab/rectilinear/controlInputRectlinearDistance} }
%     \subfigure[][]{%
%         \label{fig:weightRectilinearDistance}%
%         \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/matlab/rectilinear/weightRectilinearDistance} }
%     \caption[Leader-follower performance for rectilinear trajectory.]{First scenario (rectilinear trajectory): 
%         \subref{fig:trajectoryRectilinearDistance} leader-follower trajectories,~\subref{fig:stateErrorRectilinearDistance} state tracking errors,~\subref{fig:controlInputsRectilinearDistance} linear speed and steering angle of the follower, and    \subref{fig:weightRectilinearDistance} critic weights.}%
%     \label{fig:performanceRectilinearTrajectory}%
% \end{figure*}

 A relatively unplanned complex trajectory-tracking scenario is considered. The leader moves according to a sinusoidal trajectory so that %

%  \begin{align*}
% $x^{[\ell]}(t) = \alpha (t) \in [0,10\pi],$ %
% $y^{[\ell]}(t) = 5 \,\sin(\alpha t),$ and %
% $\theta^{[\ell]}(t) = cos(\alpha t).$  %
%        \end{align*}
% 
 The leader robot  was initially placed at $(x,y) = (0, 0)~\si{[\meter]}$ with an orientation of $\theta = 0 \degree$ while the follower robot was initially placed at a random position around $(x,y) = (0, 0)~\si{[\meter]}$ with a random orientation. The simulation results are summarized in Figure~\ref{fig:performanceSinewaveTrajectory}. It is observed that the critic weights take more time to converge compared to the earlier scenario,  which is reluctant to the complexity of the independent trajectory of the leader. This result emphasizes the adaptability of the proposed adaptive learning mechanism to different scenarios. Further, the follower starts to move away from the leader at the beginning of the simulation before it finally converges to the leader where the tracking errors are bounded by a safe distance $d.$ %

% \begin{figure*}[htbp]%
%     \centering
%     \parbox[c]{0.48\textwidth}{\centering Parallel parking}
%     \parbox[c]{0.48\textwidth}{\centering Trajectory tracking}
%     \\
%     \subfigure[][]{%
%         \label{fig:trajectorySinewave}%
%         \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/matlab/sinewave/trajectorySinewaveDistance} }
%     \subfigure[][]{%
%         \label{fig:stateErrorSinewave}%
%         \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/matlab/sinewave/stateErrorSinewaveDistance} }
%     \\
%     \subfigure[][]{%
%         \label{fig:controlInputSinewave}%
%         \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/matlab/sinewave/controlInputSinewaveDistance} }
%     \subfigure[][]{%
%         \label{fig:weightSinewaveDistance}%
%         \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/matlab/sinewave/weightSinewaveDistance} }
%     \caption[Leader-follower performance for nonlinear trajectory.]{Second scenario (sinusoidal trajectory): 
%         \subref{fig:trajectorySinewave} leader-follower trajectories,~\subref{fig:stateErrorSinewave} state tracking errors,~\subref{fig:controlInputSinewave} linear speed and steering angle of the follower, and    \subref{fig:weightSinewaveDistance} critic weights.}%
%     \label{fig:performanceSinewaveTrajectory}%
% \end{figure*}

In the third case study we rely on completley random path for the leader robot to evaluate the effectivenss of the proposed algorithm as a solution for the leader follower tracking problem. The leader is to follow a random path defined by the model described in (2), with $\gamma$ being a random value, The follwer is to maitain an offset distance of $d = 0.5 $ [m]. Note that the path of the leader is not pre defined and the algorithm being completlety model free. The results are summerized in Fig. \ref{fig:peformanceRandomTrajectory}. %

>>>>

 \section{Computer Experiments and Results}
 \label{sec:resultsExperiments}
 This section adopts the theoritical results discussed in the previous section by simulating the Algorithm using the Pioneer 3-DX in CoppeliaSim. The results obtained by the authors in \hl{cite flairs paper here} are validated using the commercial robot simulator, CoppeliaSim as a prelminary step to implement the Algorithm experimantally in real world. we present the dynamics of the tracking error and the convergence characterisitcs of the actor and critic weights. The weighting matrices are set to $\mathbf{Q} = \mathrm{diag}[0.001,0.001]$ and $\mathbf{R} = \mathrm{diag}[10^{-5}, 10^{-5}].$ % 
  \[Q=0.001 
  \begin{bmatrix}
  1       & 0   \\
  0       & 1   \\
  \end{bmatrix},
  R=0.00001
  \begin{bmatrix}
  1       & 0 \\
  0       & 1 
  \end{bmatrix}.\]
 The actor and critic learning rates $\ell_c$ are set to  $0.01$ and $0.0001$. The sampling time $T_s$ is set to $0.001 \sec.$ The desired distance offset between the leader and the follower is set to $ d = 0.5$ [m] for all scenarios.
 The CoppeliaSim simulator accuratly mimic what would happen in a realworld scenario following idustry standatrds and guidlines. The Pioneer robot model and the  was initally placed at position (-1,-2.5) m.The Target modeled as a cylinder was initially positioned on a random position around the origin (0,0). The simulation was run for 
\begin{figure}
   \centering
   \fcolorbox{gray!10}{gray!5}{
     \includegraphics[width=0.45\textwidth]{figs/coppeliasim.png}
`     }
   \caption{CoppeliaSim Scene.}
   \label{fig:CoppeliaSim Scene}
 \end{figure}

 \begin{figure*}[htbp]%
 \subfigure[][]{%
    \label{fig:trajectoryRandom}%
    \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/coppelia/firstScenario/trajectoryRandomPathDistance.eps} }
\subfigure[][]{%
    \label{fig:stateErrorRandom}%
    \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/coppelia/firstScenario/stateErrorRandomPathDistance.eps} }
\\
 \subfigure[][]{%
    \label{fig:controlInputsRandom}%
    \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/coppelia/firstScenario/controlInputRandomPathDistance.eps} }
 \subfigure[][]{%
    \label{fig:weightRandomDistance}%
    \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/coppelia/firstScenario/weightRandomPathDistance.eps} }
\caption[Leader-follower performance for random trajectory.]{Performance in tracking random trajectory: 
  \subref{fig:trajectoryRandom} leader-follower trajectories,~\subref{fig:stateErrorRandom} state error,~\subref{fig:controlInputsRandom} linear speed and steering angle of the follower; and    \subref{fig:weightRandomDistance} learning weights.}%
  \label{fig:peformanceRandomTrajectory}%
\end{figure*}

% \textit{In the sequel, computer experiments  are conducted to validate the performance of the proposed model-free adaptive learning algorithm in real-time. The results of computer experiments highlight the dynamics of the tracking errors and the convergence characteristics of the proposed algorithm (i.e., updating the critic weights). This will judge the ability of the follower robot to track independent motion trajectory of the leader robot under two different independent leader-motion trajectories. The computer experiments are realized using MATLAB simulation environment. The weighting matrices are set to $\mathbf{Q} = \mathrm{diag}[0.01,0.01,0.005]$ and $\mathbf{R} = \mathrm{diag}[10^{-6}, 10^{-6}].$ % 
%  \[Q=0.001 
%  \begin{bmatrix}
%  10       & 0   & 0\\
%  0       & 10   & 0\\
%  0       & 0   &5 \\
%  \end{bmatrix},
%  R=0.000001
%  \begin{bmatrix}
%  1       & 0 \\
%  0       & 1 
%  \end{bmatrix}.\]
% The learning rate $\ell_c$ is set to  $0.0001$. The sampling time $T_s$ is set to $0.08 \sec.$ The desired distance offset between the leader and the follower is set to $ d = 0.5$ [m] for all scenarios.
%
%% In the first s\begin{figure*}[htbp]%
%\centering
%\parbox[c]{0.48\textwidth}{\centering Parallel parking}
%\parbox[c]{0.48\textwidth}{\centering Trajectory tracking}
%\\
%%\begin{align*}
%% $\dot x^{[\ell]}(t) = \nu^{[\ell]}(t)\cos\theta^{[\ell]}(t),$ %
%% $\dot y^{[\ell]}(t) = \nu^{[\ell]}(t)\sin\theta^{[\ell]}(t).$
%% \dot\theta^\ell(t) &= \omega(t),
%%\end{align*}
% Initially, the leader's position is set to be the same for a short period of time with $\nu^{\ell} = 0 , \gamma^{\ell} = 0$. Then it starts to move on a horizontal line with $\nu^{\ell} = 0.1 $ [m/s], eventually the leader starts to move on an inclined path with $\nu^{\ell} = 0.2$ [m/s], $\gamma^{\ell} = 45\degree$. The leader is initially placed at $(x,y) = (5, 2)~\si{[\meter]}$ with an orientation of $\theta = 0 \degree $, while the follower  is initially placed at $(x,y) = (0, 0)~\si{[\meter]}$ with an orientation of $\theta = 0\degree$. Note that during this scenario, the critic weights converge rapidly and the tracking errors decay till the follower tracks the leader. The trajectory phase plan plot, tracking error states, control signals, and tuning of critic weights are shown in Fig.~\ref{fig:performanceRectilinearTrajectory}. %
%% \begin{figure*}[htbp]%
%%     \centering
%%     \parbox[c]{0.48\textwidth}{\centering Parallel parking}
%%     \parbox[c]{0.48\textwidth}{\centering Trajectory tracking}
%%     \\
%%     \subfigure[][]{%
%%         \label{fig:trajectoryRectilinearDistance}%
%%         \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/matlab/rectilinear/trajectoryRectlinearDistance} }
%%     \subfigure[][]{%
%%         \label{fig:stateErrorRectilinearDistance}%
%%         \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/matlab/rectilinear/stateErrorRectlinearDistance} }
%%     \\
%%     \subfigure[][]{%
%%         \label{fig:controlInputsRectilinearDistance}%
%%         \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/matlab/rectilinear/controlInputRectlinearDistance} }
%%     \subfigure[][]{%
%%         \label{fig:weightRectilinearDistance}%
%%         \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/matlab/rectilinear/weightRectilinearDistance} }
%%     \caption[Leader-follower performance for rectilinear trajectory.]{First scenario (rectilinear trajectory): 
%%         \subref{fig:trajectoryRectilinearDistance} leader-follower trajectories,~\subref{fig:stateErrorRectilinearDistance} state tracking errors,~\subref{fig:controlInputsRectilinearDistance} linear speed and steering angle of the follower, and    \subref{fig:weightRectilinearDistance} critic weights.}%
%%     \label{fig:performanceRectilinearTrajectory}%
%% \end{figure*}
% A relatively unplanned complex trajectory-tracking scenario is considered. The leader moves according to a sinusoidal trajectory so that %
%%  \begin{align*}
%% $x^{[\ell]}(t) = \alpha (t) \in [0,10\pi],$ %
%% $y^{[\ell]}(t) = 5 \,\sin(\alpha t),$ and %
%% $\theta^{[\ell]}(t) = cos(\alpha t).$  %
%%        \end{align*}
%% 
% The leader robot  was initially placed at $(x,y) = (0, 0)~\si{[\meter]}$ with an orientation of $\theta = 0 \degree$ while the follower robot was initially placed at a random position around $(x,y) = (0, 0)~\si{[\meter]}$ with a random orientation. The simulation results are summarized in Figure~\ref{fig:performanceSinewaveTrajectory}. It is observed that the critic weights take more time to converge compared to the earlier scenario,  which is reluctant to the complexity of the independent trajectory of the leader. This result emphasizes the adaptability of the proposed adaptive learning mechanism to different scenarios. Further, the follower starts to move away from the leader at the beginning of the simulation before it finally converges to the leader where the tracking errors are bounded by a safe distance $d.$ % 
%% \begin{figure*}[htbp]%
%%     \centering
%%     \parbox[c]{0.48\textwidth}{\centering Parallel parking}
%%     \parbox[c]{0.48\textwidth}{\centering Trajectory tracking}
%%     \\
%%     \subfigure[][]{%
%%         \label{fig:trajectorySinewave}%
%%         \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/matlab/sinewave/trajectorySinewaveDistance} }
%%     \subfigure[][]{%
%%         \label{fig:stateErrorSinewave}%
%%         \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/matlab/sinewave/stateErrorSinewaveDistance} }
%%     \\
%%     \subfigure[][]{%
%%         \label{fig:controlInputSinewave}%
%%         \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/matlab/sinewave/controlInputSinewaveDistance} }
%%     \subfigure[][]{%
%%         \label{fig:weightSinewaveDistance}%
%%         \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/matlab/sinewave/weightSinewaveDistance} }
%%     \caption[Leader-follower performance for nonlinear trajectory.]{Second scenario (sinusoidal trajectory): 
%%         \subref{fig:trajectorySinewave} leader-follower trajectories,~\subref{fig:stateErrorSinewave} state tracking errors,~\subref{fig:controlInputSinewave} linear speed and steering angle of the follower, and    \subref{fig:weightSinewaveDistance} critic weights.}%
%%     \label{fig:performanceSinewaveTrajectory}%
%% \end{figure*}
%\textit{In the third case study we rely on completley random path for the leader robot to evaluate the effectivenss of the proposed algorithm as a solution for the leader follower tracking problem. The leader is to follow a random path defined by the model described in (2), with $\gamma$ being a random value, The follwer is to maitain an offset distance of $d = 0.5 $ [m]. Note that the path of the leader is not pre defined and the algorithm being completlety model free. The results are summerized in Fig. \ref{fig:peformanceRandomTrajectory}. }
